Robots Showing Emotions: Emotion Representation with no bio-inspired Body

### 2016-10-18 JournalInteractionStudies/Parts/conclusions.tex
Addition of todos in the files, addition of a Rebuttal2.txt to start writing changes done in the text and make the rebuttal to the Editors.

Four case studies were done to study the contribution of angular velocity, linear velocity, oscillation angle, and shape change in emotion expression from a non-bio-inspired robot. The results collected during all case studies show that a correct combination of linear and angular velocity, oscillation angle, and direction can be used to convey emotions that could be distinguished from each other. Also, they suggest that it is necessary to modify the robot's shape to increase people's emotion recognition rate. This result is congruent with the results obtained in the studies done to understand what features are more relevant when humans convey emotions~\cite{Roether2009,Venture2014}.\\
The lessons learned after these four case studies can be summarized as:
\begin{itemize}
	\item It is possible to convey emotions through changes in angular and linear velocity of a non-bio-inspired body also without changes in the body shape. This lesson opens the door to the inclusion of simple "emotion" to robotic platforms that are not required to have bio-inspired characteristics, but have to interact with humans.
	\item The change in the body shape increases the emotion identification rate of the observers. However, the characteristics that should be changed to increment the emotions perception are still to be investigated.
	\item Although Laban's notation is adequate to code people's movements, robots need precise values to define their actions. As a consequence, the Laban's theory needs to be instantiated to the robots' situation, to enable comparison between different works.
	\item Emotions come as a reaction to an event. Therefore, just showing uncorrelated movements to the participants is not going to support the study of the expression of all the emotions. For example, the action Fear was well perceived by the participants when the robot get far from them, as they were inducing it.
	\item Giving the participants the possibility to write their interpretation of the robot's movements was not going to be helpful to assess what emotion they thought the robot was conveying during case studies. Although there are plenty of procedures that have been proposed to determine the emotion that a person could be thinking at (e.g., SAM), they could just be used in a controlled environment where the experimenter could spend long periods of time with the participant. This is not the case of exhibitions where people would agree to stop just for a couple of minutes. On the other side, exhibitions make it possible to collect a large number of subjects representative of a varied population.
	\item Using a real platform to perform the case studies increases the enthusiasms of the participants and possibly their willingness to participate in the study. Although we did not evaluate directly this item during our case studies, it was observed during the set up that people were curious to see the robot moving. Also, during the pilot two kids had a different reaction when the robot was getting close to them. One of them was scared of the robot, while the other was curious about the robot.
\end{itemize}
Further work should be done to understand how the combination of changes in the body shape could contribute to the expression of emotions. It is also necessary to recreate a simple scene to give the context to the participants, but this time just using two robots. This will eliminate possible clues that could be given by the human actor.

>>>>

Four case studies were done to study the contribution of angular velocity, linear velocity, oscillation angle, and shape change in emotion expression from a non-bio-inspired robot. The results collected during all case studies show that a correct combination of linear and angular velocity, oscillation angle, and direction can be used to convey emotions that could be distinguished from each other. Also, they suggest that it is necessary to modify the robot's shape to increase people's emotion recognition rate. This result is congruent with the results obtained in the studies done to understand what features are more relevant when humans convey emotions~\cite{Roether2009,Venture2014}.

The lessons learned after these four case studies can be summarized as:

\begin{itemize}

	\item It is possible to convey emotions through changes in angular and linear velocity of a non-bio-inspired body also without changes in the body shape. This lesson opens the door to the inclusion of simple "emotion" to robotic platforms that are not required to have bio-inspired characteristics, but have to interact with humans.

	\item The change in the body shape increases the emotion identification rate of the observers. However, the characteristics that should be changed to increment the emotions perception are still to be investigated.

	\item Although Laban's notation is adequate to code people's movements, robots need precise values to define their actions. As a consequence, the Laban's theory needs to be instantiated to the robots' situation, to enable comparison between different works.

	\item Emotions come as a reaction to an event. Therefore, just showing uncorrelated movements to the participants is not going to support the study of the expression of all the emotions. For example, the action Fear was well perceived by the participants when the robot get far from them, as they were inducing it.

	\item Giving the participants the possibility to write their interpretation of the robot's movements was not going to be helpful to assess what emotion they thought the robot was conveying during case studies. Although there are plenty of procedures that have been proposed to determine the emotion that a person could be thinking at (e.g., SAM), they could just be used in a controlled environment where the experimenter could spend long periods of time with the participant. This is not the case of exhibitions where people would agree to stop just for a couple of minutes. On the other side, exhibitions make it possible to collect a large number of subjects representative of a varied population.

	\item Using a real platform to perform the case studies increases the enthusiasms of the participants and possibly their willingness to participate in the study. Although we did not evaluate directly this item during our case studies, it was observed during the set up that people were curious to see the robot moving. Also, during the pilot two kids had a different reaction when the robot was getting close to them. One of them was scared of the robot, while the other was curious about the robot.

\end{itemize}

Further work should be done to understand how the combination of changes in the body shape could contribute to the expression of emotions. It is also necessary to recreate a simple scene to give the context to the participants, but this time just using two robots. This will eliminate possible clues that could be given by the human actor.

### 2016-10-18 JournalInteractionStudies/Parts/introduction.tex
Addition of todos in the files, addition of a Rebuttal2.txt to start writing changes done in the text and make the rebuttal to the Editors.

The expression of emotion in robots is becoming a quite important issue to implement robots able to interact with people in a social context. Following the results obtained by studying emotion expression in humans (e.g., \cite{Venture2014,Ekman2004}), it is possible to notice that most of the works focus on facial expressions (e.g.,~\cite{Breazeal2002}) or on humanoid poses (e.g.,~\cite{Canamero2010}), while the vast majority of robots on the market nowadays are not provided with face, arms and legs. Moreover, the focus is often on the position of limbs, or face elements, more than on the kind of movement that could express emotion, resulting on emotion expressions that could not be broadly used in most platforms.
There are applications where the shape of the robot is constrained by functional needs, such as in the case of home cleaners, lawn mowers, transporters, flying drones, and many others. In all these cases, emotions can be expressed with the only available body, exploiting movement features.
\\
The work presented in this paper focuses on the identification of movement features that could be considered platform independent, thus could be used to express emotions with a wide range of robots. We devised a very simple robot base, with no resemblance to humans or animals, and we proposed its movements to subjects, to identify how they could interpret the movements they could see from a robot in front of them. We performed live trials to keep into consideration factors that could influence the evaluation: the full perception of movement and noise, and the interaction with a physical robot are different if the same scene is lived in first person, or seen on a screen as a movie.
\\
The obtained results enable to say that it is possible to design movements, even for a neutral platform, that could be interpreted as emotional expression, with some differences among emotions, as confirmed also by other experiments~\cite{Sharma2013}.
\\
Finally, we could identify a set of very basic movement features that could be used to express emotions and, most importantly, the range of values for these features that could led to successful implementations.
\\
This paper is organized as follows. The next section provides a brief overview of relevant work related to emotion expression in robots. Section system outlines the platform used in the case studies. The section case studies presents the design and results for all four case studies.

>>>>

The expression of emotion in robots is becoming a quite important issue to implement robots able to interact with people in a social context. Following the results obtained by studying emotion expression in humans (e.g., \cite{Venture2014,Ekman2004}), it is possible to notice that most of the works focus on facial expressions (e.g.,~\cite{Breazeal2002}) or on humanoid poses (e.g.,~\cite{Canamero2010}), while the vast majority of robots on the market nowadays are not provided with face, arms and legs. Moreover, the focus is often on the position of limbs, or face elements, more than on the kind of movement that could express emotion, resulting on emotion expressions that could not be broadly used in most platforms.
There are applications where the shape of the robot is constrained by functional needs, such as in the case of home cleaners, lawn mowers, transporters, flying drones, and many others. In all these cases, emotions can be expressed with the only available body, exploiting movement features.

The work presented in this paper focuses on the identification of movement features that could be considered platform independent, thus could be used to express emotions with a wide range of robots. We devised a very simple robot base, with no resemblance to humans or animals, and we proposed its movements to subjects, to identify how they could interpret the movements they could see from a robot in front of them. We performed live trials to keep into consideration factors that could influence the evaluation: the full perception of movement and noise, and the interaction with a physical robot are different if the same scene is lived in first person, or seen on a screen as a movie.

The obtained results enable to say that it is possible to design movements, even for a neutral platform, that could be interpreted as emotional expression, with some differences among emotions, as confirmed also by other experiments~\cite{Sharma2013}.

Finally, we could identify a set of very basic movement features that could be used to express emotions and, most importantly, the range of values for these features that could led to successful implementations.

This paper is organized as follows. The next section provides a brief overview of relevant work related to emotion expression in robots. Section system outlines the platform used in the case studies. The section case studies presents the design and results for all four case studies.

### 2016-10-18 JournalInteractionStudies/Parts/system.tex
Addition of todos in the files, addition of a Rebuttal2.txt to start writing changes done in the text and make the rebuttal to the Editors.

The system used in the case studies here reported has been developed to support the identification of features that could be used to project emotions with a non-bio-inspired platform. This system is composed of two complementary parts: robotic platform and emotional enrichment system. The platform has been envisioned to be as simple as possible without any resemblance to human-like embodiment, while the emotional enrichment has been designed to be applied also to other platforms and to be extended to other emotion expressions. 
\subsection{Robotic Platform}
The platform is a holonomic base, as shown in Figure~\ref{fig:holonomic-platform}. The first prototype included three metal gear motors with encoders, an Arduino Mega micro-controller, one servo motor and X-Bee system to enable the communication between Arduino and a remote computer. The mechanical design of the prototype is shown in Figure~\ref{fig:triskar-prototype}. A PID controller for each wheel was implemented and tuned to guarantee linear and angular velocity.
\\
This prototype was modified after the first experiments to improve the expressiveness of the upper part. Two additional servo motors and a mechanical structure to support all these motors were added. The two motors attached to beams could be controlled to obtain an asymmetric or opening-closing movement. 
The third motor deforms the upper part of the foam. Figure~\ref{fig:triskar-first-design} shows the placement of the motors. To hide the structure, foam and a light blue cloth covering the foam were used, as it could be seen in Figure~\ref{fig:triskar-cover}. The light blue color was selected because it is not expected to generate any particular arousal in people~\cite{Promarker}. 
\\
Although this version showed the possibility to convey emotions, still it was not autonomous enough to host the emotional enrichment system. Therefore, an Odroid-U3 microcomputer was added, communicating with Arduino Mega via ROS-Serial.
Also, it was decided to make some improvements in the upper part. Two reasons motivated this changes. First the deformation system was not changing the shape enough to be clearly perceived by subjects; second, the beams were damaging the foam. Therefore, a lever  was added to the deformation motor, and a piece of rubber was adopted to safely connect the beams with the foam. The final arrangement is shown in Figure~\ref{fig:triskar-second-design}.\\ 
The final version, with improved computational power (Arduino Due instead of Mega) is shown in Figure~\ref{fig:triskar-third-design}.
\begin{figure}
	\centering
	\begin{subfigure}[c]{0.3\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./Images/TriskarThird.png}
	\caption{Holonomic Base used in all the platform's version.}
	\label{fig:holonomic-platform}
	\end{subfigure}
	\begin{subfigure}[c]{0.3\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./Images/upperFirstB.png}
	\caption{Platform prototype.}
	\label{fig:triskar-prototype}
	\end{subfigure}
	\begin{subfigure}[c]{0.3\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./Images/upperSecondC.png}
	\caption{First version design.}
	\label{fig:triskar-first-design}
	\end{subfigure}
	\\
	\begin{subfigure}[c]{0.3\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./Images/upperThirdD.png}
	\caption{Second version design.}
	\label{fig:triskar-second-design}
	\end{subfigure}
	\begin{subfigure}[c]{0.3\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./Images/upperFourthD.png}
	\caption{Third version design.}
	\label{fig:triskar-third-design}
	\end{subfigure}
	\begin{subfigure}[c]{0.3\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./Images/Triskar2.jpg}
	\caption{Platform with cover.}
	\label{fig:triskar-cover}
	\end{subfigure}
	\caption{Triskarino platform.}
	\label{fig:robot}
\end{figure}

>>>>

The system used in the case studies here reported has been developed to support the identification of features that could be used to project emotions with a non-bio-inspired platform. This system is composed of two complementary parts: robotic platform and emotional enrichment system. The platform has been envisioned to be as simple as possible without any resemblance to human-like embodiment, while the emotional enrichment has been designed to be applied also to other platforms and to be extended to other emotion expressions.

\subsection{Robotic Platform}
The platform is a holonomic base, as shown in Figure~\ref{fig:holonomic-platform}. The first prototype included three metal gear motors with encoders, an Arduino Mega micro-controller, one servo motor and X-Bee system to enable the communication between Arduino and a remote computer. The mechanical design of the prototype is shown in Figure~\ref{fig:triskar-prototype}. A PID controller for each wheel was implemented and tuned to guarantee linear and angular velocity.

This prototype was modified after the first experiments to improve the expressiveness of the upper part. Two additional servo motors and a mechanical structure to support all these motors were added. The two motors attached to beams could be controlled to obtain an asymmetric or opening-closing movement. 
The third motor deforms the upper part of the foam. Figure~\ref{fig:triskar-first-design} shows the placement of the motors. To hide the structure, foam and a light blue cloth covering the foam were used, as it could be seen in Figure~\ref{fig:triskar-cover}. The light blue color was selected because it is not expected to generate any particular arousal in people~\cite{Promarker}.

Although this version showed the possibility to convey emotions, still it was not autonomous enough to host the emotional enrichment system. Therefore, an Odroid-U3 microcomputer was added, communicating with Arduino Mega via ROS-Serial.
Also, it was decided to make some improvements in the upper part. Two reasons motivated this changes. First the deformation system was not changing the shape enough to be clearly perceived by subjects; second, the beams were damaging the foam. Therefore, a lever  was added to the deformation motor, and a piece of rubber was adopted to safely connect the beams with the foam. The final arrangement is shown in Figure~\ref{fig:triskar-second-design}.

The final version, with improved computational power (Arduino Due instead of Mega) is shown in Figure~\ref{fig:triskar-third-design}.
\begin{figure}[h]
	\centering
	\begin{subfigure}[c]{0.3\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./Images/TriskarThird.png}
	\caption{Holonomic Base used in all the platform's version.}
	\label{fig:holonomic-platform}
	\end{subfigure}
	\begin{subfigure}[c]{0.3\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./Images/upperFirstB.png}
	\caption{Platform prototype.}
	\label{fig:triskar-prototype}
	\end{subfigure}
	\begin{subfigure}[c]{0.3\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./Images/upperSecondC.png}
	\caption{First version design.}
	\label{fig:triskar-first-design}
	\end{subfigure}
	\\
	\begin{subfigure}[c]{0.3\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./Images/upperThirdD.png}
	\caption{Second version design.}
	\label{fig:triskar-second-design}
	\end{subfigure}
	\begin{subfigure}[c]{0.3\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./Images/upperFourthD.png}
	\caption{Third version design.}
	\label{fig:triskar-third-design}
	\end{subfigure}
	\begin{subfigure}[c]{0.3\textwidth}
	\centering
	\includegraphics[width=\textwidth]{./Images/Triskar2.jpg}
	\caption{Platform with cover.}
	\label{fig:triskar-cover}
	\end{subfigure}
	\caption{Triskarino platform.}
	\label{fig:robot}
\end{figure}
